Gas exchange in large rivers controlled by largest turbulent eddies: implications for remotely sensing gas exchange via SWOT

### 2021-07-26 manuscript_v4.Rmd
rewriting

##Key Points
- In large rivers, there exists a local equilibrium between production and dissipation of turbulent kinetic energy at the free surface
- Gas exchange in large rivers is equivalently parameterized across multiple turbulence scales in open-channel flows
- BIKER algorithm exploits this theory, predicting gas exchange velocity and $CO_2$ fluxes from simulated SWOT satellite data and in situ $CO_2$ data

@@

##Abstract
Go BIKER!!
*AGU Advances. So 250 words here*
*AGU advances is 8000 words total, though the format of the ones I've looked at vary quite a bit from long letter to short article. Many are clearly shorter than 8000 I think*

##Plain Language Summary
Necessary for AGU advances (200 words)

##1 Introduction
Since the publication of @colePlumbingGlobalCarbon2007a, the global river network is recognized to emit substantial amounts of carbon in addition to exporting to the oceans. Current estimates of total carbon dioxide ($CO_2$) evasion to the atmosphere from the global river network vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a], with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Despite its incredibly small percentage of the global land surface, this is on par with the total oceanic $CO_2$ uptake rate [@gruberOceanicSinkAnthropogenic2019] and the global forest carbon uptake rate [@panLargePersistentCarbon2011]. Though this flux is still relatively poorly constrained, it plays a critical role in the global carbon cycle and is of great interst to aquatic biogeochemists. Equation 1 represents this flux for any sparingly soluble gas given the gas concentration gradient between the water $C_{water}$ and the air $C_{air}$ and the gas exchange velocity *k*.

$$\mathbf{(1)} Flux=k(C_{water}-C_{air})$$

The structure of equation 1 necessitates that calculations of this flux are highly sensitive to one's measurements/estimates of *k*. More broadly, *k* is also of significant importanceto aquatic ecologists modeling stream metabolism [@bernhardtMetabolicRegimesFlowing2018], water quality engineers [@chapraSurfaceWaterQualityModeling2008], and general interfacial mass transport [@lorkeUnifiedScalingRelation2006]. Decades of work have focused on elucidating the physical mechanisms behind *k* in oceans and lakes and, to a lesser extent, fluvial gas exchange [see @wanninkhofAdvancesQuantifyingAirSea2009 and @hallGasExchangeStreams2020 for reviews on gas exchange in in oceans and rivers, respectively]. Given that *k* should scale with turbulence in a fluid flow, a generalized model for *k* (equation 2) relies on the characteristic turbulence velocity scale $v_K$, which itself is a function of the characteristic length and time scales [@katulStructureFunctionModel2018]. Extensive field and labratory experiments have confirmed the validity of this model when $v_K$ is set to the Kolmogorov velocity scale [e.g. @lamontEddyCellModel1970; @zappaEnvironmentalTurbulentMixing2007; @lorkeUnifiedScalingRelation2006; @zappaVariationSurfaceTurbulence2003 **MORE**], where the idea is that the turbulent kinetic energy (TKE) cascade necessitates that TKE is produced by the largest, integral-scale turbulent eddies in the flow and then passed down the cascade to progressively smaller eddies until it is dissipated as heat at the Kolmogorov-scale: the smallest eddies in a turbulent flow. Equation 2 via Kolmogorov-scale eddies can also be derived from both classic thin-film and surface renewal theories for gas exchange [@wangPhysicallyBasedScaling2021].
$$\mathbf{(2)} k\sim Sc^{-n}v_K$$

While this model works reasonably well in non-fluvial environments, there is considerable uncertainity in how it applies to fluvial systems. Specifically, the influence of the riverbed and its large roughness elements in small streams (i.e. rapids) complicates scaling fluvial *k* and might lead to substantially higher *k* in streams with whitewater [e.g. @hallAirWaterOxygen2012; @ulsethDistinctAirWater2019a]. Conversely, even less attention has been paid to the other side of the stream-to-river continumn, where large rivers' hydraulic properties begin to mimic small lakes and estuaries. To date, the handful of existing field studies of large-river *k* have suggested that their *k* properties begin to mimic small lakes with respect to wind dynamics, though little else is well-established [@alinPhysicalControlsCarbon2011; @beaulieuControlsGasTransfer2012a; @wangPhysicallyBasedScaling2021]. Further, these mechanistic uncertainties are additionally limited by a large dearth of field-measured fluvial *k*. @wangPhysicallyBasedScaling2021 attempted to address this by simulating *k* in 35 rivers using stream metabolism modeling [@applingOvercomingEquifinalityLeveraging2018] and in situ dissolved oxygen (DO) datasets, though this is still limited by the biases inherent in the DO data. This is all exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, where little in situ information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

These mechanistic uncertainties are then propogated through upscaling workflows when biogeochemists predict *k* across thousands of rivers [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] using equation 3 coupled with hydraulic geometry (HG), or the scaling relationships between streamflow and river channel hydraulics [@leopoldHydraulicGeometryStream1953a]. It is currently not well understood how sensitive global estimates of fluvial gas evasion are to the specific hydraulic geometry model that is employed by the worker. Further, these approaches rely on either in situ discharge records or modeled streamflow which introduces either additional uncertainity or the problems associated with ungauged rivers.

A potential alternative to this upscaling approach is to directly estimate a river's hydraulic properties from remote sensing (RS) data. Remote sensing of river hydraulics is a burgeoning subfield within remote sensing of hydrology, often in service of remote sensing of river discharge (RSQ- see @gleasonRemoteSensingRiver2020 for a recent review). This is accomplished via two general approaches: 'gauged' methods which rely on in situ river data to calibrate one's method to the river(s) at hand and 'ungauged' techniques which focus on generalizability and global implementation in the service of merely improving existing knowledge in data-poor domains [@gleasonRemoteSensingRiver2020]. Many, but not all, of these ungauged approaches were developed in the context of the upcoming NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) satellite mission. SWOT is expected to launch in 2022 and provide the world's first global measurements of water surface extent and elevation at novel temporal resolutions. SWOT is a wide swath radar interferometer and will sample rivers every 1 to 7 days per 21 day repeat cycle. This yields an average sampling resolution of 11 days. It will measure rivers wider than 100m with a goal of expanding this to rivers at least 50m wide [@biancamariaSWOTMissionIts2016]. A decade of SWOT work has explored the multi-parameter problem of estimating the river hydraulic parameters of roughness and bathymetry from remote sensing to produce the SWOT discharge product [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020a; @durandEstimatingReachaveragedDischarge2014a; @garamboisInferenceEffectiveRiver2015; @gleasonRetrievalRiverDischarge2014a; @garamboisVariationalEstimationEffective2020; @brissetAssimilationAltimetricData2018; @hagemannBAMBayesianAMHGManning2017a; @larnierRiverDischargeBathymetry2020a; @oubanasDischargeEstimationUngauged2018].

In this context, here we revist the fundamental mechanisms behind gas exchange in large rivers using a simple TKE budget balance and one of the largest available datasets of field-measured river hydraulics [@brinkerhoffReconcilingAtaStationAtManyStations2019] to answer the following question: does *k* behave fundamentally differently in large rivers versus other streams and how does this relate to the classical model used to explain gas evasion in non-fluvial aquatic systems (equation 2, section 3)? We exploit the findings from this simple analysis to develop a novel methodology that predicts *k* and its explicit uncertainity solely using SWOT observations (section 4). It requires no in situ inputs of any kind (although in situ data can be ingested and will improve results). We name the RS of $k_{600}$ algorithm BIKER, or the 'Bayesian Inference/Inversion of the $k_{600}$ Evasion Rate' and validate it for 47 SWOT-observable rivers from around the world using hydraulic models to produce SWOT-like data (as SWOT has not yet launched- section 2). We also quantify BIKER's sensitivity to the expected SWOT measurement errors on 17 of those rivers: while SWOT data represent a sea change in inland water monitoring, it is expected to have an approximately 10cm error in water surface elevation [@biancamariaSWOTMissionIts2016] as well as river width errors [@frassonExploringFactorsControlling2021]. Finally, we use previously published dissolved $CO_2$ data to represent a hypotetical in situ $CO_2$ sensor and compare the bulk carbon efflux from the 47 rivers as calculated using BIKER and previously published in situ techniques for predicting $k_{600}$.

@@

![Figure 1: Map of the 47 hydraulic models and 1 timeseries of CO2 samples used in this study. Note that hydraulic model locations are approximate as most of the models are not geo-referenced. We additionally used over 530,000 discrete measurements of river channel hydraulics from across the continetal United States (Brinkerhoff et al. 2019) and over 700 gas exchange velocity measurements (Ulseth et al. 2019) that are not mapped here.](cache/maps/map_fin.jpg)

TKE and gas exchange theory (section 2) are explored using two datasets. First, we use a previously published compilation of field measurements that were originally made to calibrate United States Geological Survey (USGS) streamgauge rating curves and collected by @brinkerhoffReconcilingAtaStationAtManyStations2019. That dataset contains over 530,000 unique measurements of river channel velocity, width, depth, and discharge from across the continental United States. This is to our knowledge the largest collection of field measurements of river channel hydraulics and serves as the best dataset to calculate water surface TKE budgets. Second, we use the dataset from [@ulsethDistinctAirWater2019a] who measured and/or collected (from the literature) over 700 measurements of stream hydraulics and *k* and is, to our knowledge, the largest collection of field-measured river and stream *k*. This is also the dataset used to validate our *k* predictive models.

BIKER validation (section 4) was performed on 47 SWOT-simulated rivers. Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on "SWOT-like" data. These simulated rivers are simply river-reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as RS observations and are used as the sole inputs to BIKER. Here, we use 47/51 rivers collected by @frassonExploringFactorsControlling2021 and @durandIntercomparisonRemoteSensing2016b. These are the two benchmarking studies that have explored RSQ algorithm performance for the SWOT mission. The approximate locations of these rivers are plotted in Figure 1, and please consult both of those papers for all of the hydraulic model specifications. Ultimately, the 47 rivers are spread across the United States, France, Italy, the United Kingdom, and Bangladesh. We omit three models from @durandIntercomparisonRemoteSensing2016b because they lack enough hydraulic information to calculate $U_\star$, which is necessary for algorithm validation (both Saint Lawrence River models and the Tanana River model).

@@

##3 The TKE budget in large rivers and implications for scaling river gas exchange velocity
Turbulent flow is fundamental to fluvial geomorphology and open-channel flow and has been the subject of extensive research across scientific and engineering disciplines for over a century. Turbulent kinetic energy (TKE) is the energy generated and dissipated by the eddies created from the chaotic nature of turbulent flow. @nakagawaTurbulenceOpenChannel1993 outline the standard energy budget for TKE, at some depth *h*, for a 2-dimensional open-channel flow (equation 3). *G* is the TKE produced, $\epsilon$ is the TKE dissipated, $T_d$ is TKE transported from the riverbed via turbulent diffusion, and $V_d$ is TKE transported via viscous diffusion. See Appendix A for all variable descriptions and notation.
$$\mathbf{(3)} G+T_d+V_d=\epsilon$$

$\epsilon$ is frequently described as $\epsilon(h)=\frac{U_*^3}{h}$. However in rivers, which exhibit additional depth-scale shear due to downslope flow over 'macroroughness elements' (channel geometry, bars, meanders, etc.), energy dissipation occurs much more evenly over the flow depth than this parameterization allows for. It has been shown that, for rivers at least, a more appropriate dissipation model considers stream power (per unit weight water: $\epsilon_D = gSU$), which represents the energy dissipated by the total frictional resistance of the river reach's geometry [@moogStreamReaerationNonuniform1999]. This is often called 'energy dissipation via form-drag' in the literature, and is generally used to upscale *k* in rivers and streams [i.e. @raymondScalingGasTransfer2012a; @ulsethDistinctAirWater2019a]. All three of these studies have confirmed that $\epsilon_D$ is far greater than $\epsilon$ in rivers, therefore generating the majority of fluvial TKE dissipation.

We express *G* at the free surface in two different ways (equation 4) following @nakagawaTurbulenceOpenChannel1993 on the left and our derivation (detailed in Text S1) on the right. Looking at equation 4, it is readily apparent that $\frac{R_h}{H}$ will cancel out if they are equal. This occurs in most large and wide rivers and is a common assumption in hydraulic modeling. In this scenario, the following three statements must be true: 1) $G \approx gSU \approx \epsilon_D \approx \epsilon$, 2) $\epsilon(H) \neq \frac{U_*^3}{H}$, and 3) $T_d \approx 0$. If this "large river" condition is met, then additonal TKE transported from the riverbed is functionally zero and the TKE budget at the free surface is just stream power (per unit weight water). It also implies a local equilibrium bewteen TKE produced and TKE dissipated at the free surface. 
$$\mathbf{(4)} G=\frac{UU_*^2}{H}=gSU\frac{R_h}{H} $$

We test this hypothesis using the data from @brinkerhoffReconcilingAtaStationAtManyStations2019 and @ulsethDistinctAirWater2019a by plotting boxplots of $\frac{R_h}{H}$ ratios for over 531,000 different measurements of river hydraulics versus $\epsilon_D - G$. *G* is calculated as $\frac{UU_*^2}{H}$. Note that this term is equivalent to $T_d$ in equation 1 assuming that $V_d$ is neglibile at the free surface [@nakagawaTurbulenceOpenChannel1993]. In order to calculate when $R_h=H$, we round all ratios to 3 significant figures and set all ratios greater than or equal to 0.995 as $R_h=H$. If our hypothesis holds, then $\epsilon_D - G$ should be functionally zero only when $R_h=H$.

![Figure 2: Hydraulic radius/mean depth ratios (Rh/H) versus stream power per unit weight water - TKE production (eD-G, or equivalently, Td at the free surface) for our two datasets of field-measured river channel hydraulics. eD-G is functionally zero only when the Rh = H. This indicates that, when a river is large enough, the TKE energy budget is simply governed by stream power and TKE produced is equal to TKE dissipated. A local equilibrium of TKE is acheived, where energy generated by large-scale eddies is nearly equal to energy dissipated by the smallest scale eddies. In this figure, Rh/H bins contain between 92 and 138,307 measurements. ](cache/k600_theory/turbulence.jpg)

Figure 2, which plots this test, shows that $T_d$, or $\epsilon_D-G$, is generally three to four orders of magnitude less when $R_h=H$ than it is when $\frac{R_h}{H}$ is between 90-99% (median $T_d$ of  $10^{-7.29}$ versus $10^{3.84} \frac{J}{kg*s}$, respectively). When the $\frac{R_h}{H}$ ratio is very small (0-9%), this rises to a median value of $10^{-2.56} \frac{J}{kg*s}$. We deem the $\epsilon_D - G$ values when $R_h=H$ (which are generally $10^{-6}$ to $10^{-8} \frac{J}{kg*s}$) to be functionally zero. This indicates that, when a river is large enough, the TKE energy budget at the free surface is simply stream power per unit weight water and a local equilibrium of free-surface TKE is acheived, where energy produced via the largest scale (turbulent) eddies is equivalent to energy dissipated via the smallest scale (Kolmorogov) eddies. This idea of a local equilibrium is certainly not new, however this is to our knowledge the first confirmation that it happens naturally in most large rivers. From here on we refer to the $R_h=H$ river scenario as the 'large river' condition.

This finding as a significant impact on how we model *k* in these large rivers. Because the local equilibrium acheived during the 'large river condition' necessitates that $\epsilon$ is equivalent to *G* (Figure 2), we can presumably set $v_K$ equal to the largest ('integral-scale') eddies in the system, which are much more easily remotely-sensible than the Kolmogorov velocity scale. Otherwise, *k* should continue to scale only using the Kolmogorov-scale (which is similar across all turbulent flows by its definition).

To test this hypothesis, we fit the following three models to the @ulsethDistinctAirWater2019a dataset of *K* measurements all normalized to a Schmidt number of 600: $k_{600}=\beta U_*$, $k_{600}=\alpha \epsilon^{1/4}$, and $k_{600}=a (\epsilon_D)^b$ under two scenarios: $R_h=H$ and $R_h =/= H$. Note that because all data are normalized to a Schmidt number of 600, both the Schmidt numbers and viscosity terms are constant and are implicit in the statistical coefficient. The model structures (and the fit parameters) are detailed in Table 1. The third model tested (rows 5 and 6 of Table 1), was fit to see which slope coefficient arrives empirically by fitting the generalized scaling function to the data. Therfore, the $r2$ for these models is reflective of the log-transformed $k_{600}$ (unlike the other two models). This was the approach used by @ulsethDistinctAirWater2019a, who argue that bubble-mediated gas exchange yields steeper slopes (in high-energy streams) than 0.25.

@@

![Figure 5. a: Performance metrics by river. See Table 1 for metric definitions. Dashed lines denote scores of 1, 0, and -0.41 for r2 (section 3.2). b-d: validation timeseries for three rivers representative of good, reasonable, and poor BIKER performance. b) was randomly selected from the upper tertile of NRMSE scores, c) was randomly selected from the middle tertile, and d) from the worst tertile. Model results include the posterior means and 95% CIs.](cache/validation/validation_by_river.jpg)

>>>>

##Key Points
- In large rivers, there exists a local equilibrium between the production and dissipation of turbulent kinetic energy at the free surface
- Gas exchange in large rivers is equivalently parameterized across multiple turbulence scales, with the classical Kolmogorov model recovered empirically
- BIKER algorithm exploits this theory, predicting gas exchange velocity and $CO_2$ fluxes from simulated SWOT satellite data and in situ $CO_2$ data

@@

##Abstract
*AGU Advances. So 250 words here* <br>
*AGU advances is 8000 words total, the format seems to be a long letter so I've tried to right this in that style. Might be a bold organization but we'll see*

##Plain Language Summary
*Necessary for AGU advances (200 words)*

##1 Introduction
Natural systems play a fundamental role in the budgeting and accounting of the global carbon cycle under climate change. Since the publication of @colePlumbingGlobalCarbon2007a, the global river network is recognized to emit substantial amounts of carbon in addition to exporting to the oceans. Current estimates of total carbon dioxide evasion ($FCO_2$) to the atmosphere from the global river network vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] **maybe add Shaoda here**, with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Despite its incredibly small percentage of the global land surface [0.47%- @raymondGlobalCarbonDioxide2013a], this flux is on par with the total oceanic $CO_2$ uptake rate [@gruberOceanicSinkAnthropogenic2019; @horgbyUnexpectedLargeEvasion2019] and the global forest carbon uptake rate [@panLargePersistentCarbon2011]. It is still relatively poorly constrained but is clearly a critical component of the global carbon cycle. Equation 1 represents this flux given the gas concentration gradient between the water $CO_{2_{water}}$ and the air $CO_{2_{air}}$ and the gas exchange velocity *k*.

$$\mathbf{(1)} FCO_2=k(CO_{2_{water}}-CO_{2_{air}})$$

The structure of equation 1 necessitates that calculations of this flux are highly sensitive to one's measurements/estimates of *k* [@hallGasExchangeStreams2020]. Even more broadly, fluvial *k* is of significant importance to aquatic ecologists modeling stream metabolism [e.g. @bernhardtMetabolicRegimesFlowing2018] and water quality engineers modeling river responses to pollutant loadings [among other things- @chapraSurfaceWaterQualityModeling2008]. Decades of work have focused on elucidating the physical mechanisms behind *k* in oceans and lakes [@wanninkhofAdvancesQuantifyingAirSea2009] and, to a lesser extent, fluvial gas exchange [@hallGasExchangeStreams2020]. Given that *k* should scale with turbulence in a turbulent flow, a generalized model for *k* (equation 2) relies on the characteristic turbulence velocity scale $v_K$, which itself is a function of the characteristic length and time scales [@katulStructureFunctionModel2018]. Extensive field and labratory experiments have confirmed the validity of this model when $v_K$ is set to the Kolmogorov velocity scale [e.g. @lamontEddyCellModel1970; @zappaEnvironmentalTurbulentMixing2007; @lorkeUnifiedScalingRelation2006; @zappaVariationSurfaceTurbulence2003; @moogStreamReaerationNonuniform1999; @wangCoefficientsSmallEddy2015; @vachonRelationshipNearsurfaceTurbulence2010], where the idea is that turbulent kinetic energy (TKE) is produced by the largest, integral-scale turbulent eddies in the flow and then passed down the energy cascade to progressively smaller eddies until eddy motion is stable and TKE is dissipated as heat at the Kolmogorov-scale: the smallest eddies in a turbulent flow. Equation 2 via Kolmogorov-scale eddies can also be derived from both classic thin-film and surface renewal theories for gas exchange [@wangPhysicallyBasedScaling2021].
$$\mathbf{(2)} k\sim Sc^{-n}v_K$$

While this model works reasonably well in non-fluvial environments, there is considerable uncertainity in how it applies to fluvial systems. Specifically, the influence of the riverbed and its large roughness elements in small streams (i.e. bedforms, logjams, sediment out of suspension) complicates scaling fluvial *k* and might lead to substantially higher *k* in streams with whitewater [e.g. @hallAirWaterOxygen2012; @ulsethDistinctAirWater2019a]. Conversely, even less attention has been paid to the other side of the stream-to-river continumn, which is often hypothesized, in relation to *k*, as a hybrid condition of both fluvial and non-fluvial (i.e. lentic or estuarine) hydraulic properties. To date, the handful of existing field studies of large-river *k* have suggested that *k* begins to be notably influenced by wind dynamics, though little else is well-established [@alinPhysicalControlsCarbon2011; @beaulieuControlsGasTransfer2012a; @wangPhysicallyBasedScaling2021]. These mechanistic uncertainties are additionally limited by a large dearth of field-measured fluvial *k*. @wangPhysicallyBasedScaling2021 attempted to address this by simulating *k* in 35 rivers using stream metabolism modeling [@applingOvercomingEquifinalityLeveraging2018] and in situ dissolved oxygen (DO) datasets, finding that equation 2 is valid in their simulated rivers and that *k*~streamflow relationships are different in large and small rivers. However, they did not directly compare against *k* data in extremely small systems [@ulsethDistinctAirWater2019a] and stopped short of parsing out any hydraulic explanations for why this happens.

Further, these mechanistic uncertainties are then propogated through upscaling workflows when biogeochemists predict *k* across thousands of rivers [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] via equation 1 coupled with hydraulic geometry [HG: the scaling relationships between streamflow and river channel hydraulics- @leopoldHydraulicGeometryStream1953a], and it is currently not well understood how sensitive global estimates of fluvial gas evasion are to the specific HG model that is employed by the worker. Further, these approaches rely on either in situ discharge records or modeled streamflow which introduces additional uncertainities. This is all exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, where little in situ information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

A potential alternative to this upscaling approach is to directly estimate a river's hydraulic properties from remote sensing (RS) data. Remote sensing of river hydraulics is a burgeoning subfield within remote sensing of hydrology, often in service of remote sensing of river discharge [RSQ- @gleasonRemoteSensingRiver2020]. This is accomplished via two general approaches: 'gauged' methods which rely on in situ river data to calibrate one's method to the river(s) at hand and 'ungauged' techniques which focus on hydraulic generalizability in the service of merely improving existing knowledge in data-poor domains [@gleasonRemoteSensingRiver2020]. Many, but not all, of these ungauged approaches are developed in the context of the upcoming NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) satellite mission. SWOT is expected to launch in 2022 and provide the world's first global measurements of water surface extent and elevation at novel temporal resolutions. SWOT is a wide swath radar interferometer and will sample rivers every 1 to 7 days per 21 day repeat cycle and will measure rivers wider than 100m with a goal of expanding this to rivers at least 50m wide [@biancamariaSWOTMissionIts2016]. A decade of SWOT work has explored the multi-parameter problem of estimating the river hydraulic parameters of roughness and bathymetry from remote sensing to produce the SWOT discharge product [e.g. @andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020a; @durandEstimatingReachaveragedDischarge2014a; @garamboisInferenceEffectiveRiver2015; @gleasonRetrievalRiverDischarge2014a; @garamboisVariationalEstimationEffective2020; @brissetAssimilationAltimetricData2018; @hagemannBAMBayesianAMHGManning2017a; @larnierRiverDischargeBathymetry2020a; @oubanasDischargeEstimationUngauged2018].

In this context, here we revist the fundamental mechanisms behind gas exchange and turbulence in large rivers using one of the largest available datasets of field-measured river hydraulics [@brinkerhoffReconcilingAtaStationAtManyStations2019] to answer the following question: does *k* behave fundamentally differently in large rivers than in other streams and how does this relate to the classical model of gas evasion in aquatic systems (equation 2)? We exploit the findings from this simple analysis to develop a novel methodology that predicts $k_{600}$ (*k* normalized to a Schmidt number of 600) and its explicit uncertainity solely using SWOT observations. The method requires no in situ inputs of any kind (although in situ data can be ingested and will improve results). We name the RS of $k_{600}$ algorithm BIKER, or the 'Bayesian Inference/Inversion of the $k_{600}$ Evasion Rate' and validate it for 47 SWOT-observable rivers from around the world using hydraulic models to produce SWOT-like data (as SWOT has not yet launched). We also quantify BIKER's sensitivity to the expected SWOT measurement errors on 17 of those rivers: while SWOT data represent a sea change in inland water monitoring, it is expected to have an approximately 10cm error in water surface elevation [@biancamariaSWOTMissionIts2016] as well as river width errors [@frassonExploringFactorsControlling2021]. Finally, we use previously published dissolved $CO_2$ data to represent a hypothetical in situ $CO_2$ sensor and compare the bulk carbon efflux from the 47 rivers as calculated using BIKER and previously published in situ techniques for predicting $k_{600}$.

@@

![Figure 1: Map of the 47 hydraulic models and 1 timeseries of CO2 samples used in this study. Note that hydraulic model locations are approximate as some of the models are not geo-referenced. Not mapped here are over 530,000 discrete measurements of river channel hydraulics from across the continetal United States (Brinkerhoff et al. 2019) and over 700 gas exchange velocity measurements (Ulseth et al. 2019).](cache/maps/map_fin.jpg)

TKE and gas exchange theory (section 2) are explored using two datasets. First, we use a previously published compilation of field measurements that were originally made to calibrate United States Geological Survey (USGS) streamgauge rating curves and gathered by @brinkerhoffReconcilingAtaStationAtManyStations2019. That dataset contains over 530,000 unique measurements of river channel velocity, width, depth, and discharge from across the continental United States. This is to our knowledge the largest collection of field measurements of river channel hydraulics and serves as the best dataset to calculate water surface TKE budgets. Second, we use the dataset from @ulsethDistinctAirWater2019a who measured and/or collected from the literature over 700 measurements of stream hydraulics and $k_{600}$ and is, to our knowledge, the largest collection of field-measured river and stream $k_{600}$. This is also the dataset used to validate our *k* scaling models.

BIKER validation (section 4) was performed on 47 SWOT-simulated rivers. Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on "SWOT-like" data. These simulated rivers are simply river-reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as RS observations and are used as the sole inputs to BIKER. Here, we use 47/51 rivers collected by @frassonExploringFactorsControlling2021 and @durandIntercomparisonRemoteSensing2016b. These are the two benchmarking studies that have explored RSQ algorithm performance for the SWOT mission. The approximate locations of these rivers are plotted in Figure 1, and please consult both of those papers for all of the hydraulic model specifications. Ultimately, the 47 rivers are spread across the United States, France, Italy, the United Kingdom, and Bangladesh. We omit three models from @durandIntercomparisonRemoteSensing2016b because they lack enough hydraulic information to calculate $U_\star$, which is necessary for algorithm validation. These are both Saint Lawrence River models and the Tanana River model.

@@

##3 The TKE budget in large rivers and implications for scaling fluvial gas exchange velocity
Turbulence is fundamental to fluvial geomorphology and open-channel flows and has been the subject of extensive research across scientific and engineering disciplines for over a century. Turbulent kinetic energy (TKE) is the energy generated and dissipated by the eddies created from the chaotic nature of turbulent flow. @nakagawaTurbulenceOpenChannel1993 outline the standard energy budget for TKE, at some depth *h*, for a 2-dimensional open-channel flow (equation 3). *G* is the TKE produced, $\epsilon$ is the TKE dissipated, $T_d$ is TKE transported from the riverbed via turbulent diffusion, and $V_d$ is TKE transported via viscous diffusion. See Appendix A for all variable descriptions and notation.
$$\mathbf{(3)} G+T_d+V_d=\epsilon$$

$\epsilon$ at the free surface is frequently described as $\epsilon(H)=\frac{U_*^3}{H}$. However in rivers, which exhibit additional depth-scale shear due to downslope flow over 'macroroughness elements' (i.e. channel geometry, bars, meanders, etc.), energy dissipation occurs much more evenly over the flow depth than this parameterization allows for. It has been shown that, for rivers at least, a more appropriate dissipation model considers stream power (per unit weight water: $\epsilon_D = gSU$), which represents the energy dissipated by the total frictional resistance of the river reach's geometry [@moogStreamReaerationNonuniform1999]. This is often called 'energy dissipation via form-drag' in the literature, and is generally used to upscale *k* in rivers and streams [i.e. @raymondScalingGasTransfer2012a; @ulsethDistinctAirWater2019a]. Both @raymondScalingGasTransfer2012a and @moogStreamReaerationNonuniform1999 confirmed that $\epsilon_D$ is far greater than $\epsilon$ in open-channel flows, therefore generating the majority of fluvial TKE dissipation.

We express *G* at the free surface in two different ways (equation 4) following @nakagawaTurbulenceOpenChannel1993 on the left and our derivation on the right. Text S1 details the necessary algebra to arrive at both expressions of *G*. Looking at equation 4, it is readily apparent that $\frac{R_h}{H}$ will cancel out if they are equal. This occurs in most large and wide rivers and is a common assumption in hydraulic modeling. In this scenario, the following three statements must be true: 1) $G \approx gSU \approx \epsilon_D \approx \epsilon$, 2) $\epsilon(H) \neq \frac{U_*^3}{H}$, and 3) $T_d \approx 0$. More intuitively, if this "large river" condition is met then additonal TKE transported from the riverbed is functionally zero and the TKE budget at the free surface is just stream power (per unit weight water). This implies a local equilibrium between TKE produced and TKE dissipated at the free surface. 
$$\mathbf{(4)} G=\frac{UU_*^2}{H}=gSU\frac{R_h}{H} $$

We test this hypothesis using the data from @brinkerhoffReconcilingAtaStationAtManyStations2019 and @ulsethDistinctAirWater2019a by plotting boxplots of $\frac{R_h}{H}$ ratios for over 531,000 different measurements of river hydraulics versus $\epsilon_D - G$, which is equivalent to $T_d$ in equation 3 assuming that $V_d$ is neglibile away from the wall [@nakagawaTurbulenceOpenChannel1993]. *G* is calculated as $\frac{UU_*^2}{H}$. In order to calculate when $R_h=H$, we round all ratios to 3 significant figures and set all ratios greater than or equal to 0.995 as $R_h=H$. If our hypothesis holds, then $\epsilon_D - G$ should be functionally zero only when the 'large river' condition is met.

![Figure 2: Hydraulic radius/mean depth ratios (Rh/H) versus stream power per unit weight water - TKE production (eD-G at the free surface) for our two datasets of field-measured river channel hydraulics. eD-G is functionally zero only when a river is sufficiently large. This indicates that a local equilibrium of TKE is acheived and energy generated by large-scale eddies is nearly equal to energy dissipated by the smallest scale eddies. In this figure, Rh/H bins contain between 573 and 138,307 measurements.](cache/k600_theory/turbulence.jpg)

Figure 2, which plots this test, shows that $\epsilon_D-G$, or equivalently $T_d$, is nearly two orders of magnitude less when $R_h=H$ than it is when $\frac{R_h}{H}$ is between 90-99% (median $T_d$ across both datasets of $10^{-5.39}$ versus $10^{-3.83} \frac{J}{kg*s}$, respectively). When the $\frac{R_h}{H}$ ratio is very small (0-9%), this rises to a median value of $10^{-2.56} \frac{J}{kg*s}$, or nearly three orders of magnitude. We deem the $\epsilon_D - G$ values when $R_h=H$ (which are generally $10^{-5}$ to $10^{-8} \frac{J}{kg*s}$) to be functionally zero. This means that a local equilibrium of free-surface TKE is achieved, where energy produced via the largest scale (turbulent) eddies is equivalent to energy dissipated via the smallest scale (Kolmorogov) eddies. This idea of a local equilibrium is certainly not new, however this is to our knowledge the first confirmation that it happens naturally in most large rivers and that is arises algebraically from the TKE budget in open-channel flows. From here on we refer to the $R_h=H$ river scenario as the 'large river' condition.

This finding as a significant impact on how we model *k* in large rivers. Because the local equilibrium acheived during the 'large river condition' necessitates that $\epsilon$ is equivalent to *G* (Figure 2), we can presumably set $v_K$ equal to the largest ('integral-scale') eddies in the system, which are much more easily remotely-sensible than the Kolmogorov velocity scale. Otherwise, *k* should continue to scale only using the Kolmogorov-scale (which is consistent across all turbulent flows by its definition).

To test this hypothesis, we fit the following three models to the @ulsethDistinctAirWater2019a dataset of *k* measurements normalized to a Schmidt number of 600: $k_{600}=\beta U_*$, $k_{600}=\alpha \epsilon^{1/4}$, and $k_{600}=a (\epsilon_D)^b$ under two scenarios: $R_h=H$ and $R_h =/= H$. Note that because all data are normalized to a Schmidt number of 600, both the Schmidt numbers and viscosity terms are constant and are implicit in the statistical coefficients. The model structures (and the fit parameters) are detailed in Table 1. The third model tested (rows 5 and 6 of Table 1), was fit to see which slope coefficient is retrieved empirically by fitting the generalized scaling function to the data. Therfore, the $r2$ for these models is reflective of the log-transformed $k_{600}$ (unlike the other two models). This was the approach used by @ulsethDistinctAirWater2019a, who observed slopes steeper than 0.25 in both regimes of their breakpointed regression model.

@@

![Figure 5. a: Performance metrics by river. See Table 1 for metric definitions. Dashed lines denote scores of 0, 0.50, and 1. b-d: validation timeseries for three rivers representative of good, reasonable, and poor BIKER performance. b) was randomly selected from the upper tertile of NRMSE scores, c) was randomly selected from the middle tertile, and d) from the worst tertile. See Figure S5 for all other rivers. TO DO. Model results include the posterior means and 95% CIs.](cache/validation/validation_by_river.jpg)
